---
title: "Quality control figures"
author: "Rachel Kositsky"
date: "`r Sys.Date()`"
output: html_document
params: 
  qc_excel: "ERROR_NO_PATH"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("readr")
library("tidyr")
library("dplyr")
library("ggplot2")
library("readxl")
library("scales")
```


```{r input, echo=FALSE, warning=FALSE}
#### Read in input ####
# Keep just one analysis per submission ID
df = read_excel(path = params$qc_excel) %>% distinct(submission_id, .keep_all = T)
# Make capture pool a discrete factor variable
df$capture_pool_id = as.factor(df$capture_pool_id)
df$capture_input_amount = as.factor(df$capture_input_amount)
# Make submission ID a discrete factor variable
df$submission_id = as.factor(df$submission_id)

# Check it's all from one flowcell
if(length(table(df$flowcell)) > 1) {
  print("Found more than 1 flowcell:")
  print(table(df$flowcell))
  warning("More than one flowcell found in QC table!")
}
```

## Which panel are included?

These were the panels for these submissions:
```{r panels, echo=FALSE}
# Exit if there are no submissions
if(nrow(df) == 0) {
  print("No analyses were run for this flowcell. Exiting early.")
  knitr::knit_exit()
}

table(df$panel)
```


## Read coverage

The bar charts below graph the read pairs and RNA percentage per submission.

```{r read pairs, echo=FALSE}
flowcell_title = sprintf("Flowcell %s (n=%d)", 
                         paste(unique(df$flowcell), collapse = ","), 
                         nrow(df))
MAX_X_LABELS = 20

df_long <- df %>% arrange(demux_input_read_pairs) %>% 
  select(demux_dna_read_pairs, demux_rna_read_pairs, capture_pool_id, submission_id) %>%
  gather("Nucelic_Acid", "Value", -capture_pool_id, -submission_id)

df_x_format <- function(b) {
  if (nrow(df)/2 < MAX_X_LABELS) {
    return(b)
  } else {
    return("")
  }
}

df_long_x_format <- function(b) {
  if (nrow(df_long)/2 < MAX_X_LABELS) {
    return(b)
  } else {
    return("")
  }
}

ggplot(data = df_long,
       aes(x = submission_id, y = Value, fill = Nucelic_Acid)) + 
  geom_col() + scale_y_continuous(labels = comma) +
  scale_x_discrete(labels = df_long_x_format) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(subtitle = flowcell_title,
       title = "Read pairs per submission", y = "Read pairs")

ggplot(data = df_long, 
       aes(x = submission_id, y = Value, fill = Nucelic_Acid)) + 
  geom_bar(position = "fill", stat = "identity") +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = df_long_x_format) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(subtitle = flowcell_title,
       title = "DNA/RNA breakdown per submission", 
       y = "Percentage of reads")
```

## PCR duplicates

#### PCR duplicates colored by capture pool
For DNA:
```{r DNA PCR duplicates, echo = FALSE}
# DNA PCR dups
p <- ggplot(data = df,
       aes(x = submission_id, y = dna_pct_pcr_dups/100, fill = capture_pool_id)) + 
  geom_hline(yintercept = 0.25, linetype = "dashed", alpha = 0.5) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = df_x_format) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(subtitle = flowcell_title,
       title = "DNA PCR duplicate rate per submission",
       y = "DNA percent PCR duplicates")
if (max(df$dna_pct_pcr_dups, na.rm = T) > 50) {
  p <- p + geom_hline(yintercept = 0.50, linetype = "dashed", alpha = 0.5)
}
if (max(df$dna_pct_pcr_dups, na.rm = T) > 70) {
  p <- p + geom_hline(yintercept = 0.75, linetype = "dashed", alpha = 0.5)
}
print(p)
```

For RNA:
```{r RNA PCR duplicates, echo = FALSE}
# RNA PCR dups
p <- ggplot(data = df, 
            aes(x = submission_id, y = rna_pct_pcr_dups/100, fill = capture_pool_id)) + 
  geom_hline(yintercept = 0.25, linetype = "dashed", alpha = 0.5) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = df_x_format) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(subtitle = flowcell_title,
       title = "RNA PCR duplicate rate per submission",
       y = "RNA percent PCR duplicates")
if (max(df$rna_pct_pcr_dups, na.rm = T) > 50) {
  p <- p + geom_hline(yintercept = 0.50, linetype = "dashed", alpha = 0.5)
}
if (max(df$rna_pct_pcr_dups, na.rm = T) > 70) {
  p <- p + geom_hline(yintercept = 0.75, linetype = "dashed", alpha = 0.5)
}
print(p)
```

#### PCR duplicates colored by capture input volume
For DNA:
```{r DNA PCR duplicates capture input, echo = FALSE}
# DNA PCR dups
p <- ggplot(data = df,
       aes(x = submission_id, y = dna_pct_pcr_dups/100, fill = capture_input_volume)) + 
  geom_hline(yintercept = 0.25, linetype = "dashed", alpha = 0.5) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = df_x_format) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(subtitle = flowcell_title,
       title = "DNA PCR duplicate rate per submission",
       y = "DNA percent PCR duplicates")
if (max(df$dna_pct_pcr_dups, na.rm = T) > 50) {
  p <- p + geom_hline(yintercept = 0.50, linetype = "dashed", alpha = 0.5)
}
if (max(df$dna_pct_pcr_dups, na.rm = T) > 70) {
  p <- p + geom_hline(yintercept = 0.75, linetype = "dashed", alpha = 0.5)
}
print(p)
```

For RNA:
```{r RNA PCR duplicates capture input, echo = FALSE}
# RNA PCR dups
p <- ggplot(data = df, 
            aes(x = submission_id, y = rna_pct_pcr_dups/100, fill = capture_input_volume)) + 
  geom_hline(yintercept = 0.25, linetype = "dashed", alpha = 0.5) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = df_x_format) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(subtitle = flowcell_title,
       title = "RNA PCR duplicate rate per submission",
       y = "RNA percent PCR duplicates")
if (max(df$rna_pct_pcr_dups, na.rm = T) > 50) {
  p <- p + geom_hline(yintercept = 0.50, linetype = "dashed", alpha = 0.5)
}
if (max(df$rna_pct_pcr_dups, na.rm = T) > 70) {
  p <- p + geom_hline(yintercept = 0.75, linetype = "dashed", alpha = 0.5)
}
print(p)
```

#### PCR duplicates colored by capture input amount
For DNA:
```{r DNA PCR duplicates capture input amt, echo = FALSE}
# DNA PCR dups
p <- ggplot(data = df,
       aes(x = submission_id, y = dna_pct_pcr_dups/100, fill = capture_input_amount)) + 
  geom_hline(yintercept = 0.25, linetype = "dashed", alpha = 0.5) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = df_x_format) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(subtitle = flowcell_title,
       title = "DNA PCR duplicate rate per submission",
       y = "DNA percent PCR duplicates")
if (max(df$dna_pct_pcr_dups, na.rm = T) > 50) {
  p <- p + geom_hline(yintercept = 0.50, linetype = "dashed", alpha = 0.5)
}
if (max(df$dna_pct_pcr_dups, na.rm = T) > 70) {
  p <- p + geom_hline(yintercept = 0.75, linetype = "dashed", alpha = 0.5)
}
print(p)
```

For RNA:
```{r RNA PCR duplicates capture input amt, echo = FALSE}
# RNA PCR dups
p <- ggplot(data = df, 
            aes(x = submission_id, y = rna_pct_pcr_dups/100, fill = capture_input_amount)) + 
  geom_hline(yintercept = 0.25, linetype = "dashed", alpha = 0.5) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = df_x_format) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(subtitle = flowcell_title,
       title = "RNA PCR duplicate rate per submission",
       y = "RNA percent PCR duplicates")
if (max(df$rna_pct_pcr_dups, na.rm = T) > 50) {
  p <- p + geom_hline(yintercept = 0.50, linetype = "dashed", alpha = 0.5)
}
if (max(df$rna_pct_pcr_dups, na.rm = T) > 70) {
  p <- p + geom_hline(yintercept = 0.75, linetype = "dashed", alpha = 0.5)
}
print(p)
```


## Coverage
DNA coverage:
```{r DNA coverage, echo=FALSE}
median_cov = median(df$dna_mean_coverage)
p <- ggplot(data = df,
            aes(x = submission_id, y = dna_mean_coverage, fill = capture_pool_id)) + 
  geom_hline(yintercept = median_cov, alpha = 0.5) +
  geom_col() +
  scale_x_discrete(labels = df_x_format) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(subtitle = flowcell_title,
       title = sprintf("DNA mean coverage per submission [median %2.0fx]", median_cov),
       y = "DNA mean coverage (x)")
print(p)
```

RNA coverage:
```{r RNA coverage, echo=FALSE}
median_cov = median(df$rna_mean_coverage)
p <- ggplot(data = df,
            aes(x = submission_id, y = rna_mean_coverage, fill = capture_pool_id)) + 
  geom_hline(yintercept = median_cov, alpha = 0.5) +
  geom_col() +
  scale_x_discrete(labels = df_x_format) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(subtitle = flowcell_title,
       title = sprintf("RNA mean coverage per submission [median %2.0fx]", median_cov),
       y = "RNA mean coverage (x)")
print(p)
```

## On-target performance

DNA on-target:
```{r DNA on-target, echo=FALSE}
median_on_target = median(df$dna_pct_on_target)
ggplot(data = df,
       aes(x = submission_id, y = dna_pct_on_target/100, fill = capture_pool_id)) + 
  geom_col() +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = df_x_format) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(subtitle = flowcell_title,
       title = sprintf("DNA on-target rate per submission [median %2.0fx]", median_on_target),
       y = "DNA percent on-target")
```

RNA on-target:
```{r RNA on-target, echo=FALSE}
median_on_target = median(df$rna_pct_on_target)
ggplot(data = df,
       aes(x = submission_id, y = rna_pct_on_target/100, fill = capture_pool_id)) + 
  geom_col() +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = df_x_format) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(subtitle = flowcell_title,
       title = sprintf("RNA on-target rate per submission [median %2.0fx]", median_on_target),
       y = "RNA percent on-target")
```
